### 2018-03-19 submission/manuscript.Rmd
complete fixes for result section

***Minimal Bray-Curtis Differences are Detected by Cycle Number:***
A few small differences based on sub-sampling and cycle number were detected in overall bacterial community composition. Within the same fecal sample and independent of HiFi DNA polymerases, there were differences in the community composition between 20 and 25 cycles that was dependent on sub-sampling depth (sub-sampled to 1000 = `r round(median(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`) (median (IQR)), sub-sampled to 5000 = `r round(median(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`), sub-sampled to 10000 = `r round(median(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3A]. Further, when data was available for the mock communities, there were larger observed differences between 20 and 25 cycles (sub-sampled to 1000 = `r round(median(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3B]. Additionally, these stated community differences disappear when comparing 25 to 30 cycles and do not persist past 25 cycles [Figure 3]. Although these trends are clearly noticeable, we found that there was no detectable difference in Bray-Curtis index when comparing to the previous 5-cycle increment for both fecal and mock communities after multiple comparison correction (P-value `r ifelse(min(bray_sim_to_prev_cycle_kruskal_results$bh, na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). Using PERMANOVA to test for community differences based on cycle number within HiFi DNA polymerases, only Phusion had cycle dependent differences at 1000 and 5000 sub-sampling depths (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)` and `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 5000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`). For fecal samples, Phusion was one of two HiFi DNA polymerases that had enough sequences to reach a sub-sampling depth of 1000 at 15 cycles. Overall, these data suggest that there are small HiFi DNA polymerase differences in Bray-Curtis index that are dependent on both sub-sampling depth and cycle number.

***Sequence Error is Dependent on both HiFi DNA Polymerase and Cycle Number:***
Differences by HiFi DNA polymerase in the median average per base error varied without a clear pattern across sub-sampling depth [Table S5]. The highest per base median average error rates were for the Kappa HiFi DNA polymerase [Figure 4]. This error rate was minimally affected by both the 'pre.cluster' step and chimera removal by VSEARCH [Figure 4]. The differences in the per base error rate between the various HiFi DNA polymerase was cycle dependent with Platinum having the largest differences versus other HiFi DNA polymerases [Figure 4B-C and Table S6]. The total sequences with at least one error had was also cycle number dependent and mostly alleviated by the use of the 'pre.cluster' step [Figure S1]. Differences in sequences with an error before this 'pre.cluster' step were due to large differences in Accuprime^TM^ and Platinum versus the other HiFi DNA polymerases [Figure S1 & Table S7 & S8]. Investigation of whether there were HiFi DNA polymerase dependent effects on base substitution found that there were no biases in the types of substitution made [Figure S2]. Although though sequence error is dependent on HiFi DNA polymerase it can be corrected using exsiting bioinformatic approaches.

***Chimeric Sequences are HiFi DNA Polymerase Dependent and Correlate with Number of OTUs:***
At all levels of sub-sampling and cycle numbers there were significant differences in the chimera prevalence based on HiFi DNA polymerase used (P-value `r ifelse(max(filter(kruskal_chimera_prev, sub_sample_level > 500)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S9]. Differences between Platinum and all other HiFi DNA polymerases accounted for the majority of these differences and was independent of cycle number and sub-sampling depth [Table S10]. Across sub-sampling depth and cycle number Accuprime^TM^ had the lowest chimera prevalence of all the HiFi DNA Polymerases regardless of whether 'pre.cluster' or chimera removal with VSEARCH was used [Figure 5].

For all HiFi DNA polymerases, a positive correlation was observed between chimeric sequences and number of OTUs, with this correlation being strongest for Accuprime, Platinum and Phusion HiFi DNA Polymerase [Figure 6]. The R^2^ value between the number of OTUs and chimeric sequences did not change with the use of 'pre.cluster' or with the removal of chimeras using VSEARCH [Figure 6]. Taken together, this data suggests that a strong correlation exists between the number of OTUs and the prevalence of chimeric sequences.

>>>>

***Minimal Bray-Curtis Differences are Detected and are Dependent on both Cycle Number and Sub-Sampling Depth:***
A few small differences based on sub-sampling and cycle number were detected in overall bacterial community composition. Within the same fecal sample and independent of HiFi DNA polymerases, there were differences in the community composition between 20 and 25 cycles that was dependent on sub-sampling depth (sub-sampled to 1000 = `r round(median(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`) (median (IQR)), sub-sampled to 5000 = `r round(median(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["5000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`), sub-sampled to 10000 = `r round(median(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(dist_data[["10000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3A]. Further, when data was available for the mock communities, there were larger observed differences between 20 and 25 cycles (sub-sampled to 1000 = `r round(median(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"]), digits = 2)` (`r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["25%"], digits = 2)` - `r round(quantile(filter(mock_dist_data[["1000"]], cycle_compare == "20to25")[, "distance"])["75%"], digits = 2)`)) [Figure 3B]. Additionally, these stated community differences disappear when comparing 25 to 30 cycles and do not persist past 25 cycles [Figure 3]. Although these trends are clearly noticeable, we found that there was no detectable difference in Bray-Curtis index when comparing to the previous 5-cycle increment for both fecal and mock communities after multiple comparison correction (P-value `r ifelse(min(bray_sim_to_prev_cycle_kruskal_results$bh, na.rm = T) > 0.05, invisible(">"), invisible("check data"))` 0.05). Using PERMANOVA to test for community differences based on cycle number within HiFi DNA polymerases, only Phusion had cycle dependent differences at 1000 and 5000 sub-sampling depths (P-value = `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 1000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)` and `r format(filter(bray_permanova_by_taq_results, sub_sample_level == 5000, taq == "PHU")[, "bh"], digits = 1, nsmall = 1)`). For fecal samples, Phusion was one of two HiFi DNA polymerases that had enough sequences to reach a sub-sampling depth of 1000 at 15 cycles. Overall, these data suggest that there are small HiFi DNA polymerase differences in Bray-Curtis index that are dependent on both sub-sampling depth and cycle number.

***Sequence Error is Dependent on both HiFi DNA Polymerase and Cycle Number:***
Differences in the median average per base error varied by HiFi DNA polymerase without a clear pattern across sub-sampling depth [Table S5]. The highest median average per base error rates were for the Kappa HiFi DNA polymerase [Figure 4]. This error rate was minimally affected by both the 'pre.cluster' step and chimera removal by VSEARCH [Figure 4]. The differences in the median average per base error rate between the different HiFi DNA polymerases was cycle dependent with Platinum having the largest changes versus other HiFi DNA polymerases [Figure 4B-C and Table S6]. The total sequences with at least one error was also cycle number dependent and differences between HiFi DNA polymerases could be drastically reduced by the use of the 'pre.cluster' step [Figure S1]. These differences in sequences with at least one error were mostly due to differences in Accuprime^TM^ and Platinum versus the other HiFi DNA polymerases [Figure S1 & Table S7 & S8]. Finally, we did not observe a HiFi DNA polymerase dependent difference on base substitution rate [Figure S2]. Although sequence error is dependent on HiFi DNA polymerase some of these error dependent differences can be corrected using exsiting bioinformatic approaches.

***Prevalence of Chimeric Sequences are HiFi DNA Polymerase Dependent and Correlate with the Number of OTUs:***
There were significant differences in the chimera prevalence based on HiFi DNA polymerase used at all levels of sub-sampling and cycle numbers (P-value `r ifelse(max(filter(kruskal_chimera_prev, sub_sample_level > 500)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S9]. Differences in chimera prevalence between Platinum and all other HiFi DNA polymerases accounted for the majority of these differences [Table S10]. Accuprime^TM^ had the lowest chimera prevalence of all HiFi DNA polymerases regardless of whether 'pre.cluster' or chimera removal with VSEARCH was used [Figure 5]. A positive correlation was observed between chimeric sequences and the number of OTUs for all HiFi DNA polymerases [Figure 6]. This positive correlation was strongest for Accuprime^TM^, Platinum, and Phusion HiFi DNA Polymerases [Figure 6]. The R^2^ value between the number of OTUs and chimeric sequences did not change with the use of 'pre.cluster' or with the removal of chimeras using VSEARCH [Figure 6]. This data suggests that chimera prevalence depends on HiFi DNA polymerase used and confirms that the number of OTUs is dependent on the prevalence of these chimeric sequences.

### 2018-03-19 submission/manuscript.tex
complete fixes for result section

\textbf{\emph{Minimal Bray-Curtis Differences are Detected by Cycle
Number:}} A few small differences based on sub-sampling and cycle number
were detected in overall bacterial community composition. Within the
same fecal sample and independent of HiFi DNA polymerases, there were
differences in the community composition between 20 and 25 cycles that
was dependent on sub-sampling depth (sub-sampled to 1000 = 0.51 (0.4 -
0.79) (median (IQR)), sub-sampled to 5000 = 0.43 (0.33 - 0.63),
sub-sampled to 10000 = 0.4 (0.24 - 0.43)) {[}Figure 3A{]}. Further, when
data was available for the mock communities, there were larger observed
differences between 20 and 25 cycles (sub-sampled to 1000 = 0.88 (0.42 -
0.91)) {[}Figure 3B{]}. Additionally, these stated community differences
disappear when comparing 25 to 30 cycles and do not persist past 25
cycles {[}Figure 3{]}. Although these trends are clearly noticeable, we
found that there was no detectable difference in Bray-Curtis index when
comparing to the previous 5-cycle increment for both fecal and mock
communities after multiple comparison correction (P-value \textgreater{}
0.05). Using PERMANOVA to test for community differences based on cycle
number within HiFi DNA polymerases, only Phusion had cycle dependent
differences at 1000 and 5000 sub-sampling depths (P-value = 0.03 and
0.01). For fecal samples, Phusion was one of two HiFi DNA polymerases
that had enough sequences to reach a sub-sampling depth of 1000 at 15
cycles. Overall, these data suggest that there are small HiFi DNA
polymerase differences in Bray-Curtis index that are dependent on both
sub-sampling depth and cycle number.

\textbf{\emph{Sequence Error is Dependent on both HiFi DNA Polymerase
and Cycle Number:}} Differences by HiFi DNA polymerase in the median
average per base error varied without a clear pattern across
sub-sampling depth {[}Table S5{]}. The highest per base median average
error rates were for the Kappa HiFi DNA polymerase {[}Figure 4{]}. This
error rate was minimally affected by both the `pre.cluster' step and
chimera removal by VSEARCH {[}Figure 4{]}. The differences in the per
base error rate between the various HiFi DNA polymerase was cycle
dependent with Platinum having the largest differences versus other HiFi
DNA polymerases {[}Figure 4B-C and Table S6{]}. The total sequences with
at least one error had was also cycle number dependent and mostly
alleviated by the use of the `pre.cluster' step {[}Figure S1{]}.
Differences in sequences with an error before this `pre.cluster' step
were due to large differences in Accuprime\textsuperscript{TM} and
Platinum versus the other HiFi DNA polymerases {[}Figure S1 \& Table S7
\& S8{]}. Investigation of whether there were HiFi DNA polymerase
dependent effects on base substitution found that there were no biases
in the types of substitution made {[}Figure S2{]}. Although though
sequence error is dependent on HiFi DNA polymerase it can be corrected
using exsiting bioinformatic approaches.

\textbf{\emph{Chimeric Sequences are HiFi DNA Polymerase Dependent and
Correlate with Number of OTUs:}} At all levels of sub-sampling and cycle
numbers there were significant differences in the chimera prevalence
based on HiFi DNA polymerase used (P-value \textless{} 0.05) {[}Table
S9{]}. Differences between Platinum and all other HiFi DNA polymerases
accounted for the majority of these differences and was independent of
cycle number and sub-sampling depth {[}Table S10{]}. Across sub-sampling
depth and cycle number Accuprime\textsuperscript{TM} had the lowest
chimera prevalence of all the HiFi DNA Polymerases regardless of whether
`pre.cluster' or chimera removal with VSEARCH was used {[}Figure 5{]}.

For all HiFi DNA polymerases, a positive correlation was observed
between chimeric sequences and number of OTUs, with this correlation
being strongest for Accuprime, Platinum and Phusion HiFi DNA Polymerase
{[}Figure 6{]}. The R\textsuperscript{2} value between the number of
OTUs and chimeric sequences did not change with the use of `pre.cluster'
or with the removal of chimeras using VSEARCH {[}Figure 6{]}. Taken
together, this data suggests that a strong correlation exists between
the number of OTUs and the prevalence of chimeric sequences.

>>>>

\textbf{\emph{Minimal Bray-Curtis Differences are Detected and are
Dependent on both Cycle Number and Sub-Sampling Depth:}} A few small
differences based on sub-sampling and cycle number were detected in
overall bacterial community composition. Within the same fecal sample
and independent of HiFi DNA polymerases, there were differences in the
community composition between 20 and 25 cycles that was dependent on
sub-sampling depth (sub-sampled to 1000 = 0.51 (0.4 - 0.79) (median
(IQR)), sub-sampled to 5000 = 0.43 (0.33 - 0.63), sub-sampled to 10000 =
0.4 (0.24 - 0.43)) {[}Figure 3A{]}. Further, when data was available for
the mock communities, there were larger observed differences between 20
and 25 cycles (sub-sampled to 1000 = 0.88 (0.42 - 0.91)) {[}Figure
3B{]}. Additionally, these stated community differences disappear when
comparing 25 to 30 cycles and do not persist past 25 cycles {[}Figure
3{]}. Although these trends are clearly noticeable, we found that there
was no detectable difference in Bray-Curtis index when comparing to the
previous 5-cycle increment for both fecal and mock communities after
multiple comparison correction (P-value \textgreater{} 0.05). Using
PERMANOVA to test for community differences based on cycle number within
HiFi DNA polymerases, only Phusion had cycle dependent differences at
1000 and 5000 sub-sampling depths (P-value = 0.03 and 0.01). For fecal
samples, Phusion was one of two HiFi DNA polymerases that had enough
sequences to reach a sub-sampling depth of 1000 at 15 cycles. Overall,
these data suggest that there are small HiFi DNA polymerase differences
in Bray-Curtis index that are dependent on both sub-sampling depth and
cycle number.

\textbf{\emph{Sequence Error is Dependent on both HiFi DNA Polymerase
and Cycle Number:}} Differences in the median average per base error
varied by HiFi DNA polymerase without a clear pattern across
sub-sampling depth {[}Table S5{]}. The highest median average per base
error rates were for the Kappa HiFi DNA polymerase {[}Figure 4{]}. This
error rate was minimally affected by both the `pre.cluster' step and
chimera removal by VSEARCH {[}Figure 4{]}. The differences in the median
average per base error rate between the different HiFi DNA polymerases
was cycle dependent with Platinum having the largest changes versus
other HiFi DNA polymerases {[}Figure 4B-C and Table S6{]}. The total
sequences with at least one error was also cycle number dependent and
differences between HiFi DNA polymerases could be drastically reduced by
the use of the `pre.cluster' step {[}Figure S1{]}. These differences in
sequences with at least one error were mostly due to differences in
Accuprime\textsuperscript{TM} and Platinum versus the other HiFi DNA
polymerases {[}Figure S1 \& Table S7 \& S8{]}. Finally, we did not
observe a HiFi DNA polymerase dependent difference on base substitution
rate {[}Figure S2{]}. Although sequence error is dependent on HiFi DNA
polymerase some of these error dependent differences can be corrected
using exsiting bioinformatic approaches.

\textbf{\emph{Prevalence of Chimeric Sequences are HiFi DNA Polymerase
Dependent and Correlate with the Number of OTUs:}} There were
significant differences in the chimera prevalence based on HiFi DNA
polymerase used at all levels of sub-sampling and cycle numbers (P-value
\textless{} 0.05) {[}Table S9{]}. Differences in chimera prevalence
between Platinum and all other HiFi DNA polymerases accounted for the
majority of these differences {[}Table S10{]}.
Accuprime\textsuperscript{TM} had the lowest chimera prevalence of all
HiFi DNA polymerases regardless of whether `pre.cluster' or chimera
removal with VSEARCH was used {[}Figure 5{]}. A positive correlation was
observed between chimeric sequences and the number of OTUs for all HiFi
DNA polymerases {[}Figure 6{]}. This positive correlation was strongest
for Accuprime\textsuperscript{TM}, Platinum, and Phusion HiFi DNA
Polymerases {[}Figure 6{]}. The R\textsuperscript{2} value between the
number of OTUs and chimeric sequences did not change with the use of
`pre.cluster' or with the removal of chimeras using VSEARCH {[}Figure
6{]}. This data suggests that chimera prevalence depends on HiFi DNA
polymerase used and confirms that the number of OTUs is dependent on the
prevalence of these chimeric sequences.
